Group Theory in Physics (Course Note)

### 2016-11-14 Lecture note - Group Theory in Physics/sections/point-group.tex
Add a useful formula for reflection and rotation.

\paragraph{Conjugacy Classes}
Note that by fact \ref{fact:20161010-srs} one can see that such
difference mentioned will make the two classes of $2$-fold axises in
even $n$ each form their own conjugacy class. On the contrary, by the
same fact \ref{fact:20161010-srs} one sees that when $n$ is odd, all
the $2$-fold axises are in the same conjugacy class. Note that a
reflection is its own inverse, so in either cases ($n$ odd or even),
we have found self-conjugate classes.

@@

\paragraph{Formal definition or $D_n$} Having discussed about
conjugacy classes, one sees that $D_n$ can be defined by generators.
Clearly, when $n$ is odd, $D_n$ is generated by one rotation $2\pi/n$
and one reflection (denoted $C'_2$ in the book \cite{book}). When $n$
is even, one might thinks that we have three generators, i.e. the
rotation by $2\pi/n$, one reflection about vortex $C'_2$, and one
reflection about middle of edges $C''_2$. But actually it only needs
two. The reflection about middle of edges can be produced by the
rotation and the reflection about vortex, or vice versa. See this
note for example:

@@

\paragraph{Invariant Subgroups} (\textbf{pp.30 of \cite{book}})
By the distinction between edge and vortices type of $2$-fold axises
discussed before, one can see that there is difference in invariant
subgroup when $n$ is odd or even. For example, when $n=5$ is odd,
$D_5$ has only $C_5$ as a notrivial subgroup. But when
$n=6$ is even, it has $C_6$ (then $C_2$ and $C_3$), two copies of
$D_3$ (each formed by edge and vortex type of $2$-fold axis).

@@

\paragraph{Character Table} (\textbf{pp.66 of \cite{book}})

@@

\paragraph{Irreducible representation} (\textbf{pp.71-74 of \cite{book}})

>>>>

\subsubsection{Conjugacy Classes}
Note that by fact \ref{fact:20161010-srs} one can see that such
difference mentioned will make the two classes of $2$-fold axises in
even $n$ each form their own conjugacy class. On the contrary, by the
same fact \ref{fact:20161010-srs} one sees that when $n$ is odd, all
the $2$-fold axises are in the same conjugacy class. Note that a
reflection is its own inverse, so in either cases ($n$ odd or even),
we have found self-conjugate classes.

@@

\subsubsection{Formal definition of \texorpdfstring{$D_n$}{}}

Having discussed about conjugacy classes, one sees that $D_n$ can be
defined by generators.  Clearly, when $n$ is odd, $D_n$ is generated
by one rotation $2\pi/n$ and one reflection (denoted $C'_2$ in the
book \cite{book}). When $n$ is even, one might thinks that we have
three generators, i.e. the rotation by $2\pi/n$, one reflection about
vortex $C'_2$, and one reflection about middle of edges $C''_2$. But
actually it only needs two. The reflection about middle of edges can
be produced by the rotation and the reflection about vortex, or vice
versa. See this note for example:

@@

Observe that by this definition, it's easy to prove that
\begin{align}
    SR &= R^{-1}S \\
    SR^{-1} &= RS
\end{align}
which are all physically intuitive. A more sophisticated law is
also physically easy to see
\begin{fact}
    \begin{equation}
        S_m R^n = R^{-n}S_m
    \end{equation}
\end{fact}
But I here give a mathematical derivation of the above formula:
\begin{proof}
    Assume $n$ is add.

    Pick a reference point in the $n$-polygon, label its vortex as $0$. Then
    let $S_0= S$ be the reflection about this vortex axis. Then label
    all the remaining vortex $2$-fold axises in the same direction as
    $R$ as $1,2,\cdots,n-1$. Obviously we have:
    \begin{align}
        S_1 &= RS_0 R^{-1} = RSR^{-1} \nonumber\\
        S_2 &= RS_1 R^{-1} = R^2 S R^{-2} \nonumber\\
        \cdots & \nonumber \\
        S_{m} &= R^{m} S R^{m}
    \end{align}
    where $m=1,2,\cdots,n-1$. Then
    \begin{align*}
        S_m R^n &= R^m S R^{-m} R^n \\
          &= R^m S R^n R^{-m}\\
          &= R^m R^{-n} S R^{-m} \\
          &= R^{-n} S_m
    \end{align*}

    When $n$ is odd, the proof is the same, except that now $S_{2m}$ is a
    vortex reflection, and $S_{2m+1}$ is an edge reflection
    ($m=0,1,\cdots (n/2-1)$). Notice that $S_1= RS_0 = RS$, we easily
    see
    \begin{align}
        S_{2m} &= R^m S R^{-m} \\
        S_{2m+1} &= R^{m+1} S R^{-m}
    \end{align}
    and the rest of the proof is the same.
\end{proof}

\subsubsection{Invariant Subgroups} (\textbf{pp.30 of \cite{book}})
By the distinction between edge and vortices type of $2$-fold axises
discussed before, one can see that there is difference in invariant
subgroup when $n$ is odd or even. For example, when $n=5$ is odd,
$D_5$ has only $C_5$ as a notrivial subgroup. But when
$n=6$ is even, it has $C_6$ (then $C_2$ and $C_3$), two copies of
$D_3$ (each formed by edge and vortex type of $2$-fold axis).

@@

\subsubsection{Character Table} (\textbf{pp.66 of \cite{book}})

@@

\subsubsection{Irreducible representation}

(\textbf{pp.71-74 of \cite{book}})
